---
title: bsvars
author: Rob J Hyndman
date: 27 June 2025
toc: true
format:
  presentation-beamer:
    pdf-engine: pdflatex
    knitr:
      opts_chunk:
        dev: "CairoPDF"
    keep-tex: true
    fig-width: 8
    fig-height: 4.3
    template-partials:
      - before-title.tex
highlight-style: tango
execute:
  echo: false
  message: false
  warning: false
  cache: false
---

```{r}
#| label: load-packages
library(ggplot2)
library(kableExtra)
library(bsvars)
library(bsvarSIGNs)

options(digits = 4, width = 80)
ggplot2::theme_set(
  theme_get() + theme(text = element_text(family = "Fira Sans"))
)
```

# bsvars

* Appreciate focus on design
* Use generic forecast() otherwise it won't play nicely with other packages. Note horizon rather than h as second argument.
* Add print method for Forecasts
* plots can't be easily modified. Use ggplot2

```{r}
library(bsvars)
# specify the model and set seed
set.seed(2025 - 6 - 27)
us_fiscal_lsuw |>
  specify_bsvar$new(p = 1) |>
  estimate(S = 5, show_progress = FALSE) |>
  estimate(S = 10, show_progress = FALSE) |>
  forecast(horizon = 4) -> predictive

predictive

data(us_fiscal_ex_forecasts) # upload exogenous variables future values
data(us_fiscal_cond_forecasts) # upload a matrix with projected ttr

us_fiscal_lsuw |>
  specify_bsvar$new(p = 1, exogenous = us_fiscal_ex) |>
  estimate(S = 5, show_progress = FALSE) |>
  estimate(S = 10, show_progress = FALSE) |>
  forecast(
    horizon = 8,
    exogenous_forecast = us_fiscal_ex_forecasts,
    conditional_forecast = us_fiscal_cond_forecasts
  ) |>
  plot(col = "#c24b14")
```

## Links

\begin{textblock}{7.6}(4.2,3.6)
\begin{alertblock}{}
\fontsize{18}{22}\sf\bfseries\centering
\href{https://robjhyndman.com/osfs25}{robjhyndman.com/osfs25}
\end{alertblock}
\end{textblock}


\begin{textblock}{7.6}(4.2,5.6)
\begin{block}{}
\fontsize{18}{22}\sf\bfseries\centering
\href{https://robjhyndman.com/postdoc}{robjhyndman.com/postdoc}
\end{block}
\end{textblock}
